Manuscript figures

### 2022-09-30 docs/manuscript_figures.Rmd
Figure 7 for revision

load(here('docs/figures/conf_mean_std.rda'))
pred_path <- here(file.path('results/dl/prediction',
                           'unet_k15_only_clr_200epc',
                           'landcover_north_withwl.tif'))
color_tb <- data.frame(
    id = 1:8,
    type = c("Cropland", "Forest/Dense tree", 
             "Grassland", "Shrubland", 
             "Water", 'Built-up', "Bareland", 
             'Wetland'),
    color = c('#ff7f00', '#074702', '#b2df8a',
              '#33a02c', '#3347f9', '#5a6057',
              '#fdbf6f', '#0ea591'))
pred <- read_stars(pred_path)
names(pred) <- 'landcover'
# pred <- rast(pred_path)
# set.cats(pred, value = color_tb$type, index = color_tb$id)
# pred <- st_as_stars(pred)

# plot
# Get examples
tiles <- read_sf('data/geoms/tiles_nicfi_north.geojson')
id_selected <- c('1226-1012', '1227-1005', '1209-997', '1219-1017',
                 '1234-992', '1211-1008', '1236-1007','1220-995')
tiles <- tiles %>% filter(tile %in% id_selected) %>% 
    st_transform(crs = crs(rast(pred_path))) %>% 
    mutate(id = 1:nrow(.))
rst_tiles <- lapply(1:nrow(tiles), function(n){
    tm <- st_crop(pred, tiles %>% slice(n)) %>% 
        tm_shape() + 
        tm_raster("landcover", 
              n = 8, style = 'cat',
              labels = color_tb$type,
              palette = color_tb$color) +
        tm_layout(frame = F,
                  legend.show = F,
                  title.position = c(0, 0.9),
                  title = tiles %>% slice(n) %>% pull(id))
     tmap_grob(tm)
})
plot_left <- ggarrange(
    rst_tiles[[1]], rst_tiles[[2]], rst_tiles[[3]], 
    rst_tiles[[4]], ncol = 1, nrow = 4)
plot_right <- ggarrange(
    rst_tiles[[5]], rst_tiles[[6]], rst_tiles[[7]],
    rst_tiles[[8]], ncol = 1, nrow = 4)

tmap_mode("plot")
pred_final <- tm_shape(pred) +
    tm_raster("landcover", 
              title = "Land cover types",
              n = 8, style = 'cat',
              labels = color_tb$type,
              palette = color_tb$color) +
    tm_shape(tiles) +
    tm_borders("black", lwd = 1.2) +
    tm_text("id", col = 'black', fontface = "bold", size = 0.8) +
    tm_layout(frame = F,
              legend.show = F,
              outer.margins = c(0.02, 0.01, 0, 0.01),
              inner.margins = 0)
pred_final <- tmap_grob(pred_final)

# Get legend
r_df <- data.frame(x = 1:8, 
                   y = 1:8,
                   rst = as.factor(1:8))
legend_lc <- ggplot() + 
    geom_raster(
        data = r_df, 
        aes(x = x, y = y, 
            fill = rst)) +
    scale_fill_manual('Land cover',
                      values = color_tb$color, 
                      breaks = 1:8,
                      labels = color_tb$type) +
    theme_minimal() +
    theme(legend.position = 'right',
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          plot.margin = unit(rep(0, 4),"cm"))+
    guides(fill = guide_legend(ncol = 8))
legend_lc <- get_legend(legend_lc)

tbl_theme <- gridExtra::ttheme_default(
    core = list(fg_params=list(cex = 0.6),
                bg_params = list(fill = 'gray95')),
    colhead = list(fg_params=list(cex = 0.7)),
    rowhead = list(fg_params=list(cex = 0.7)))
tbl_conf <- tableGrob(mean_std, rows = NULL, theme = tbl_theme)
legend_lc <- ggarrange(NA, tbl_conf, NA, legend_lc, NA,
                       ncol = 5, widths = c(1, 2, 1, 2, 1))

pred_final_plt <- ggarrange(pred_final, legend_lc, NA,
                        nrow = 3, heights = c(5, 1, 0.6))
pred_example <- ggarrange(plot_left, pred_final_plt, plot_right,
          ncol = 3, widths = c(1, 3, 1))
ggsave(here('docs/figures/pred_example.png'),
       pred_example,
       width = 7, height = 6, dpi = 300)
```

@@

# Confusion matrix
cf_orig <- data.frame(cf$table) %>% 
    select(Prediction, Freq, Reference) %>% 
    pivot_wider(names_from = Reference, 
                values_from = Freq)
cf_orig$num <- rowSums(cf_orig[, -1], na.rm = TRUE)
cf_orig$proportion <- cls_prop$proportion
cf_ratio <- do.call(rbind, lapply(1:nrow(cf_orig), function(n){
    cf_orig[n, 2:8] / cf_orig[[n, 9]] * cf_orig[[n, 10]]
}))

@@

# 3D pie
cols <- color_tb$color
lab <- paste(round(adjust_stats_area$area, 1), "km2")
plotrix::pie3D(adjust_stats_area$proportion,
      col = cols,
      labels = lab,
      explode = 0.15,
      labelcex = 1,
      radius = 1.1,
      border = "white",
      theta = 1.1)

# Version 2 of final map
tmap_mode("plot")
pred_final <- tm_shape(pred) +
    tm_raster("landcover_north_withwl.tif", 
              #title = "Land cover types",
              title = '',
              n = 8, style = 'cat',
              labels = color_tb$type,
              palette = color_tb$color) +
    tm_layout(frame = F,
              legend.show = FALSE,
              outer.margins = c(0.02, 0.01, 0, 0.01),
              inner.margins = 0)
pred_final <- tmap_grob(pred_final)

r_df <- data.frame(x = 1:8, 
                   y = 1:8,
                   rst = as.factor(1:8))
legend_lc <- ggplot() + 
    geom_raster(
        data = r_df, 
        aes(x = x, y = y, 
            fill = rst)) +
    scale_fill_manual('',
                      values = color_tb$color, 
                      breaks = 1:8,
                      labels = color_tb$type) +
    theme_minimal() +
    theme(legend.position = 'right',
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          plot.margin = unit(rep(-1, 4),"cm"))+
    guides(fill = guide_legend(ncol = 4))
legend_lc <- get_legend(legend_lc)
f1 <- ggarrange(pred_final, legend_lc, NA, 
          nrow = 3, heights = c(5, 1, 0.5))

# Get pie chart for class frequency
cls_freq <- rast(pred_path)
cls_freq <- freq(cls_freq) %>% 
    data.frame() %>% 
    mutate(landcover = value,
           ratio = count / sum(.$count)) %>% 
    select(landcover, ratio)
save(cls_freq, file = 'docs/figures/cls_freq.rda')

cls_freq <- cls_freq %>% arrange(landcover)
cols <- color_tb$color
lab <- paste0(round(cls_freq$ratio/sum(cls_freq$ratio) * 100, 1), "%")

# Barplot
cls_freq <- cls_freq %>% 
    mutate(landcover = factor(landcover, 
                              levels = cls_freq %>% 
                               arrange(desc(ratio)) %>% 
                               pull(landcover)),
        label_ypos = 0.9 * ratio)
barplot <- ggplot(data=cls_freq, aes(x=landcover, y=ratio)) +
    geom_bar(stat="identity", aes(fill=as.factor(landcover))) +
    geom_text(aes(y=label_ypos, label=lab), angle = 90,
              color="black", size=3)+
    scale_fill_manual(values = cols[as.integer(levels(cls_freq$landcover))]) +
    theme_void() +
    theme(legend.position = "none")

@@

tbl_theme <- gridExtra::ttheme_default(
    core = list(fg_params=list(cex = 0.7, hjust = 0, x = 0.1,
                               parse = TRUE),
                bg_params = list(fill = 'gray95')),
    colhead = list(fg_params=list(cex = 0.8)),
    rowhead = list(fg_params=list(cex = 0.8)))
tbl_conf <- tableGrob(adjust_stats_area, rows = NULL, theme = tbl_theme)

pred_final_plt <- ggarrange(f1,
                            ggarrange(NA, tbl_conf,
                                      barplot, NA, 
                                      nrow = 4,
                                      labels = c(NA, 'B', 'C', NA),
                                      hjust = c(2, 4, 2, 2),
                                      heights = c(0.2, 1, 1, 0.3)), 
                            NA,
                            labels = c('A', NA, NA),
                            vjust = 5, hjust = -1,
                            ncol = 3, widths = c(6, 2, 0.5))
ggsave(here('docs/figures/final_map.png'),
       pred_final_plt,
       width = 8, height = 5.5, dpi = 300)
```

@@

fnames_s1 <- file.path("data/plts_nicfi",
                       sprintf('planet_medres_normalized_analytic_2017-12_2018-05_mosaic_%s.tif', 
                               id_selected))
fnames_s2 <- file.path("data/plts_nicfi",
                       sprintf('planet_medres_normalized_analytic_2018-06_2018-11_mosaic_%s.tif', 
                               id_selected))
fg_s1 <- lapply(fnames_s1, function(fname) {
    x <- stack(fname) %>% subset(1:3) %>% 
        raster::stretch(minq = 0.02, maxq = 0.98)
    tm <- tm_shape(x) +
        tm_rgb(r = 1, g = 2, b = 3,
               max.value = max(values(x)[, 1:3]))
    tmap_grob(tm)
    # stack(fname) %>% subset(1:4) %>% 
    #     ggRGB(r = 1, g = 2, b = 3, 
    #           scale = max(values(.)[, 1:3]), 
    #           stretch = 'lin') +
    #     theme_minimal() +
    #     theme(axis.title.x = element_blank(),
    #           axis.text.x = element_blank(),
    #           axis.ticks.x = element_blank(),
    #           axis.title.y = element_blank(),
    #           axis.text.y = element_blank(),
    #           axis.ticks.y = element_blank(),
    #           plot.margin = unit(rep(0, 4),"cm"))
})

fg_s2 <- lapply(fnames_s2, function(fname) {
    x <- stack(fname) %>% subset(1:3) %>% 
        raster::stretch(minq = 0.02, maxq = 0.98)
    tm <- tm_shape(x) +
        tm_rgb(r = 1, g = 2, b = 3,
               max.value = max(values(x)[, 1:3]))
    tmap_grob(tm)
    # stack(fname) %>% subset(1:4) %>% 
    #     ggRGB(r = 1, g = 2, b = 3, 
    #           scale = max(values(.)[, 1:3]), 
    #           stretch = 'hist') +
    #     theme_minimal() +
    #     theme(axis.title.x = element_blank(),
    #           axis.text.x = element_blank(),
    #           axis.ticks.x = element_blank(),
    #           axis.title.y = element_blank(),
    #           axis.text.y = element_blank(),
    #           axis.ticks.y = element_blank(),
    #           plot.margin = unit(rep(0, 4),"cm"))
})
s1 <- ggarrange(fg_s1[[1]], fg_s1[[2]], 
          fg_s1[[3]], fg_s1[[4]], 
          nrow = 4, ncol = 1)
s2 <- ggarrange(fg_s2[[1]], fg_s2[[2]], 
          fg_s2[[3]], fg_s2[[4]], 
          nrow = 4, ncol = 1)

>>>>

load(here('docs/figures/conf_mean_std.rda'))
pred_path <- here(file.path('results/dl/prediction',
                           'unet_k15_only_clr_200epc',
                           'landcover_north_withwl.tif'))
color_tb <- data.frame(
    id = 1:8,
    type = c("Cropland", "Forest/Dense tree", 
             "Grassland", "Shrubland", 
             "Water", 'Built-up', "Bareland", 
             'Wetland'),
    color = c('#ff7f00', '#074702', '#b2df8a',
              '#33a02c', '#3347f9', '#5a6057',
              '#fdbf6f', '#0ea591'))
pred <- read_stars(pred_path)
names(pred) <- 'landcover'

# plot
# Get examples
tiles <- read_sf('data/geoms/tiles_nicfi_north.geojson')
id_selected <- c('1211-1008', '1209-997', '1221-1012', '1234-992')
tiles <- tiles %>% filter(tile %in% id_selected) %>% 
    st_transform(crs = crs(rast(pred_path))) %>% 
    mutate(id = letters[1:4])
```

@@

# Confusion matrix
load("docs/figures/test_conf.rda")
cf_orig <- data.frame(cf$table) %>% 
    select(Prediction, Freq, Reference) %>% 
    pivot_wider(names_from = Reference, 
                values_from = Freq)
cf_orig$num <- rowSums(cf_orig[, -1], na.rm = TRUE)
cf_orig$proportion <- cls_prop$proportion
cf_ratio <- do.call(rbind, lapply(1:nrow(cf_orig), function(n){
    cf_orig[n, 2:8] / cf_orig[[n, 9]] * cf_orig[[n, 10]]
}))

@@

# Version 2 of final map
tmap_mode("plot")
pred_final <- tm_shape(pred) +
    tm_raster("landcover", 
              #title = "Land cover types",
              title = '',
              n = 8, style = 'cat',
              labels = color_tb$type,
              palette = color_tb$color) +
    tm_shape(tiles) +
    tm_borders("black", lwd = 1.2) +
    tm_text("id", col = 'black', fontface = "bold", 
            size = 1.3, xmod = 1.1,ymod = 0.1) +
    tm_layout(frame = F,
              legend.show = FALSE,
              outer.margins = rep(0, 4),
              inner.margins = 0)
pred_final <- tmap_grob(pred_final)

r_df <- data.frame(x = 1:8, 
                   y = 1:8,
                   rst = as.factor(1:8))
legend_lc <- ggplot() + 
    geom_raster(
        data = r_df, 
        aes(x = x, y = y, 
            fill = rst)) +
    geom_polygon(data = r_df, fill = NA, color = "black", size = 1,
            aes(x = x, y = y, lty = 'solid')) +
    scale_fill_manual('',
                      values = color_tb$color, 
                      breaks = 1:8,
                      labels = color_tb$type) +
    scale_linetype_identity(name = '',
                            breaks = c('solid'),
                            labels = c('Examples in Figure 7'),
                            guide = 'legend') +
    theme_void() +
    theme(legend.position = 'right',
          text = element_text(size = 14),
          legend.text = element_text(size = 14),
          plot.margin = unit(rep(-1, 4),"cm"))+
    guides(fill = guide_legend(ncol = 4))
legend_lc <- get_legend(legend_lc)
f1 <- ggarrange(pred_final, legend_lc, NA, 
                nrow = 3, heights = c(5, 1, 0.5))

# # Get pie chart for class frequency
# cls_freq <- rast(pred_path)
# cls_freq <- freq(cls_freq) %>% 
#     data.frame() %>% 
#     mutate(landcover = value,
#            ratio = count / sum(.$count)) %>% 
#     select(landcover, ratio)
# save(cls_freq, file = 'docs/figures/cls_freq.rda')
# 
# 
# cls_freq <- cls_freq %>% arrange(landcover)
# cols <- color_tb$color
# lab <- paste0(round(cls_freq$ratio/sum(cls_freq$ratio) * 100, 1), "%")
# 
# # Barplot
# cls_freq <- cls_freq %>% 
#     mutate(landcover = factor(landcover, 
#                               levels = cls_freq %>% 
#                                arrange(desc(ratio)) %>% 
#                                pull(landcover)),
#         label_ypos = 0.9 * ratio)
# barplot <- ggplot(data=cls_freq, aes(x=landcover, y=ratio)) +
#     geom_bar(stat="identity", aes(fill=as.factor(landcover))) +
#     geom_text(aes(y=label_ypos, label=lab), angle = 90,
#               color="black", size = 4)+
#     scale_fill_manual(values = cols[as.integer(levels(cls_freq$landcover))]) +
#     theme_void() +
#     theme(legend.position = "none",
#           plot.margin = unit(c(-0.1, -0.1, 0.1, -0.1),"cm"))

# Barplot
cls_freq <- data.frame(area = unlist(class_area) / whole_area * 100,
                       err = c(adjust_class_err / whole_area * 100, 0),
                       landcover = color_tb$id) %>% 
    mutate(err_adjust_prop = glue(
        paste0('{format(round(area, digits = 1), nsmall = 1)}',
               '\u00B1{format(round(err, digits = 1), nsmall = 1)}',
               '%'))) %>% 
    arrange(-area) %>% 
    mutate(label_ypos = c(0.75, 0.75, 0.52, 2.9, 3, 3.8, 5.2, 12)* area)
cls_freq <- cls_freq %>% 
    mutate(landcover = factor(landcover, 
                              levels = cls_freq %>% 
                                  arrange(desc(area)) %>% 
                                  pull(landcover)))
barplot <- ggplot(data=cls_freq, aes(x=landcover, y = area)) +
    geom_bar(stat="identity", aes(fill=as.factor(landcover))) +
    geom_errorbar(aes(ymin = area - err, ymax = area + err), 
                  width = .4, color = "black", size = 0.6,
                  position = position_dodge(.9)) +
    geom_text(aes(y=label_ypos, label=err_adjust_prop), angle = 90,
              color="black", size = 4)+
    scale_fill_manual(values = cols[as.integer(levels(cls_freq$landcover))]) +
    theme_void() +
    theme(legend.position = "none",
          plot.margin = unit(c(-0.1, -0.1, 0.1, -0.1),"cm"))

@@

tbl_theme <- gridExtra::ttheme_default(
    core = list(fg_params=list(cex = 1, hjust = 0, x = 0.1,
                               parse = TRUE),
                bg_params = list(fill = 'gray95')),
    colhead = list(fg_params=list(cex = 1)),
    rowhead = list(fg_params=list(cex = 1)))
tbl_conf <- tableGrob(adjust_stats_area, rows = NULL, theme = tbl_theme)

pred_final_plt <- ggarrange(
    f1,
    ggarrange(NA, tbl_conf,
              barplot, NA, 
              nrow = 4,
              labels = c(NA, 'B', 'C', NA),
              font.label = list(face = "plain", size = 16),
              hjust = c(2, 6, 2, 2),
              heights = c(0.1, 1, 1, 0.1)), 
    NA,
    labels = c('A', NA, NA),
    font.label = list(face = "plain", size = 16),
    vjust = 5, hjust = -1,
    ncol = 3, widths = c(6, 2, 0.5))

# Figure 7, combination art
# 1.5 column figure , 2756 * 1700 at 500 dpi
ggsave(here('docs/figures/figure7_final_map.tiff'),
       pred_final_plt, bg = "white", 
       width = 2756, height = 1700, 
       units = "px")
```

@@

fnames_s1 <- file.path(
    "data/plts_nicfi",
    sprintf('planet_medres_normalized_analytic_2017-12_2018-05_mosaic_%s.tif', 
                               id_selected))
fnames_s2 <- file.path(
    "data/plts_nicfi",
    sprintf('planet_medres_normalized_analytic_2018-06_2018-11_mosaic_%s.tif', 
            id_selected))
fg_s1 <- lapply(fnames_s1, function(fname) {
    x <- stack(fname) %>% subset(1:3) %>% 
        raster::stretch(minq = 0.02, maxq = 0.98)
    tm <- tm_shape(x) +
        tm_rgb(r = 1, g = 2, b = 3,
               max.value = max(values(x)[, 1:3]))
    tmap_grob(tm)
})

fg_s2 <- lapply(fnames_s2, function(fname) {
    x <- stack(fname) %>% subset(1:3) %>% 
        raster::stretch(minq = 0.02, maxq = 0.98)
    tm <- tm_shape(x) +
        tm_rgb(r = 1, g = 2, b = 3,
               max.value = max(values(x)[, 1:3]))
    tmap_grob(tm)
})
s1 <- ggarrange(fg_s1[[1]], fg_s1[[2]], 
          fg_s1[[3]], fg_s1[[4]], 
          nrow = 4, ncol = 1)
s2 <- ggarrange(fg_s2[[1]], fg_s2[[2]], 
          fg_s2[[3]], fg_s2[[4]], 
          nrow = 4, ncol = 1)
